Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-28 thesis/chapters/ch_chapter4.tex
Updates to chapter 4 including a new figure.

We summarize the source-averaged results for all enrichment events in runs II\_E46, II\_E49, II\_E50, II\_E51, and II\_E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, solid), CGM (black, dashed) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
\textit{statement on how immediate enrichement is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej}$ ~ 0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. This implies that, if one to were assume that all metals behaved the same and adopted an $f_{\rm ej}$ of $\sim$0.9 -- typical for the total metal ejection fraction in simulations that do not track individual enrichment channels \citep[e.g.][]{Muratov2017,Christensen2018}, and consistent with observations of the O abundance in Local Group dwarf galaxies \citep[e.g.][]{Kirby2011,McQuinn2015} -- one would underestimate the amount of AGB metals in the ISM by a factor of $\sim$ 3, and overestimate exotic, high-energy sources by a factor of $\sim$ 2. In each case, these are metals that cannot participate in the enrichment of future stellar populations.

In all cases, the metals contained with the ISM are deposited predominately in the ionized phases of the ISM, the WIM or the HIM, with the relative fraction in each phase driven by the energy of the event. Metals in the ISM injected with $E_{\rm ej} > 10^{49}$~erg are initially located predominately in the HIM, roughly 1/2 for $E_{\rm ej} = 10^{49}$~erg and nearly all for $E_{\rm ej} > 10^{52}$~erg. The lowest energy events are initially in the WIM and WNM, tracing the two dominant volume-filling components of the ISM (see Chatper~\ref{ch:chapter1}), as these events do not have sufficient energy to generate the HIM phase by themselves. In general, very few of these elements are available for immediate star formation in the CNM. However, we note that we may not have sufficient resolution to properly resolve the details of the initial mixing of individual enrichment events in the ISM. In addition, we are missing important physical processes, like dust production in AGB winds and core collapse SNe, that may affect these results for specific enrichment channels in ways that do not depend solely on $E_{\rm ej}$. Resolving these processes in detail in galaxy-scale simulations is computationally infeasible.

Finally, we are interested in the long-term evolution of the metals from each of these sources, and how long it takes for these metals to propogate from the warm / hot phases of the ISM to cold, star forming gas. Gas above the star formation threshold in our simulations is limited and short-lived due to the low gas surface densities and star formation rates in this galaxy and the efficiency of stellar feedback from newly formed stars. As a proxy, we examine the evolution of the CNM, out of which the star forming gas forms. In Figure~\ref{ch4:fig:CNM}, we examine the evolution of the metal fraction of the CNM \textit{for just those elements retained in the ISM] for the average behavior of each enrichment event.

\begin{figure*}
  \centering
  \includegraphics[width=0.75\linewidth]{figures/ch4/CNM_evolution_average}
  \caption{Time evolution of the metal fraction of the CNM for just those metals retained in the ISM for each enrichment source. In other words,
           these are the same CNM lines as shown in Figure~\ref{ch4:fig:ISM_CGM}, but are divided by the black, dashed lines in each panel.}
  \label{ch4:fig:CNM}
\end{figure*}
% make a figure here showing JUST the CNM evolution of the ISM for each source

>>>>

We summarize the source-averaged results for all enrichment events in runs II\_E46, II\_E49, II\_E50, II\_E51, and II\_E52 in Figure~\ref{ch4:fig:ISM_CGM} by showing the fraction of source metals contained within each phase of the ISM (colored lines, which sum to the black dashed line), and the CGM (black, solid). These plots show a clear, immediate trend across $E_{\rm ej}$ for all lines in the figure, but the differences are most striking for the ISM (black, solid), CGM (black, dashed) lines. In Chapter~\ref{ch:chapter3}
% \citep{Emerick2019}
we demonstrated that there was a significant difference in metal ejection fraction, $f_{\rm ej}$, for metals from AGB sources as compared to metals released in SNe. Here, we confirm that this is driven predominantly by differences in the energy of the events, and not necessarily when or where they occur in the ISM. The clear trend in Figure~\ref{ch4:fig:ISM_CGM} shows that events with higher $E_{\rm ej}$ are much more readily ejected from the disk of the galaxy, while events with lower $E_{\rm ej}$ couple poorly to the same galactic outflows.
\textit{statement on how immediate enrichment is after events}
% By 10~Myr after the enrichment events,
The high-energy events rapidly converge on a peak $f_{\rm ej}$ within about 20~Myr of the event, with only a gradual increase towards the end of the 150~Myr as the metals in the ISM are swept up in additional outflows. The lower energy events evolve more gradually. The lowest energy event, corresponding to AGB winds, reaches $f_{\rm ej}$ ~ 0.68, which increases to 0.87 for the 10$^{51}$~erg events, and 0.95 for the 10$^{52}$~erg event. To emphasize these differences, we plot each line separately in the left panel of Figure~\ref{ch4:fig:CGM_CNM}

This implies that, if one to were assume that all metals behaved the same and adopted an $f_{\rm ej}$ of $\sim$0.9 -- typical for the total metal ejection fraction in simulations that do not track individual enrichment channels \citep[e.g.][]{Muratov2017,Christensen2018}, and consistent with observations of the O abundance in Local Group dwarf galaxies \citep[e.g.][]{Kirby2011,McQuinn2015} -- one would underestimate the amount of AGB metals in the ISM by a factor of $\sim$ 3, and overestimate exotic, high-energy sources by a factor of $\sim$ 2. In each case, these are metals that cannot participate in the enrichment of future stellar populations.

In all cases, the metals contained with the ISM are deposited predominately in the ionized phases of the ISM, the WIM or the HIM, with the relative fraction in each phase driven by the energy of the event. Metals in the ISM injected with $E_{\rm ej} > 10^{49}$~erg are initially located predominately in the HIM, roughly 1/2 for $E_{\rm ej} = 10^{49}$~erg and nearly all for $E_{\rm ej} > 10^{52}$~erg. The lowest energy events are initially in the WIM and WNM, tracing the two dominant volume-filling components of the ISM (see Chapter~\ref{ch:chapter1}), as these events do not have sufficient energy to generate the HIM phase by themselves.

Gas above the star formation threshold in our simulations is limited and short-lived due to the low gas surface densities and star formation rates in this galaxy and the efficiency of stellar feedback from newly formed stars. As a proxy, we examine the evolution of the CNM, out of which the star forming gas forms. In general, very few of these elements are available for immediate star formation in the CNM ($<< 1\%$, see Figure~\ref{ch4:fig:CGM_CNM}). However, we note that we may not have sufficient resolution to properly resolve the details of the initial mixing of individual enrichment events in the ISM. In addition, this value will be sensitive to whether or not a given event occurs in the vicinity of or inside an active, star forming region -- as is likely for core collapse SNe, and unlikely for other events. Thus, this is a quantity that does not depend solely on $E_{\rm ej}$ and is sensitive to whether or not self-enrichment of star forming regions is an important factor in galactic chemical evolution. We are also missing important physical processes, like dust production in AGB winds and core collapse SNe, that may affect these results for specific enrichment channels in ways that do not depend solely on $E_{\rm ej}$. Investigating the fraction of metals immediately available for star formation will require additional, high-resolution simulations of metal mixing in and around individual star forming regions.
% add references (maybe) to high res simulation of dust production in AGB winds and SNe ???

What we can examine, however, is the final quantity of interest: the long-term evolution of the metals from each of these sources, and how long it takes for these metals to propagate from the warm / hot phases of the ISM to cold, star forming gas. In the right panel of Figure~\ref{ch4:fig:CGM_CNM}, we examine the evolution of the metal fraction of the CNM \textit{for just those elements retained in the ISM}. Although the initial CNM fractions are about the same for each source, the evolution over the first $\sim$50~Myr is qualitatively different. the higher energy sources, $E_{\rm ej} > 10^{49}$~erg, are more rapidly incorporated into the CNM than E46, even though each retains a lower fraction of metals in the ISM. This is most significant at $\sim$20~Myr, when the fraction of metals in the CNM from these sources is a factor of $\sim$3-4 higher than the E46 metals. By $\sim$50~Myr, the evolutions become intermingled and complex, with no clear trend as a function of $E_{\rm ej}$.

\begin{figure*}
  \centering
  \includegraphics[width=0.45\linewidth]{figures/ch4/CGM_average_evolution}
  \includegraphics[width=0.45\linewidth]{figures/ch4/CNM_average_evolution}
  \caption{Time evolution of the fraction of enrichment source metals contained in the CGM over time (left), and the fraction of metals in the CNM just those metals retained in the ISM for each enrichment source (right). These are the same as the lines in Figure~\ref{ch4:fig:ISM_CGM}, but the CNM lines are normalized by the black, dashed total ISM line in Figure~\ref{ch4:fig:ISM_CGM}.}
  \label{ch4:fig:CGM_CNM}
\end{figure*}
% make a figure here showing JUST the CNM evolution of the ISM for each source

However, the metals from each source do gradually trend towards a fraction of $\sim$0.8 by the end of the 150~Myr simulation time, corresponding to the mass fraction of the CNM at this point in the galaxy's evolution (see Figure~\ref{ch1:fig:}). This trend -- that the fraction of metals contained in a given phases tends towards the mass fraction of that phase -- is true across all phases in the simulation. Therefore, we can consider that the metals for each source are well-mixed across the \textit{phases} of the ISM on timescales of $\sim$100-150~Myr. \textbf{Compute the cooling timescale of each phase (maybe with a phase diagram?) and also the dynamical timescale and make comparisons here.}. These two results suggest that the delay-time between when metals are available to enrich ongoing star formation is on order of $\sim$10~Myr across sources, and only significant for the lowest (E46, AGB-like) energy sources. Although well-mixed across phases, we emphasize that this does not imply that the abundances are the same across phases nor that the metals are spatially well-mixed across the galaxy. We investigate this second point further below.
